### 2012-12-15 RSLC/paper/paper.tex
Start writeup; have abstract

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

%\documentclass[preprint]{sigplanconf}
\documentclass{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathpartir}
\usepackage{url}
\usepackage{graphics}

% ____________________________________________________________
% Listings Package Configuration
% \usepackage[scaled]{beramono}

\renewcommand*\ttdefault{txtt}
\usepackage[T1]{fontenc}

\RequirePackage{listings}
% \lstdefinelanguage{Rust}{
%   morekeywords={fn,let,enum,if,else,return,mut,const,copy,ref,move,match,struct,type,int,while,float,as,fail},
%   sensitive=true,
%   morecomment=[l]{//},
%   morecomment=[s]{/*}{*/},
%   morestring=[b]",
% }
% \lstset{language=Rust,
%         frame=none,%single,
%         basicstyle=\ttfamily\footnotesize,
%         keywordstyle=\bfseries,
%         commentstyle=\footnotesize,
%         tabsize=2,
%         showstringspaces=false,
%         flexiblecolumns=false,
%         mathescape=true,
%         escapeinside={\#}{\#},
%         % numbers=left, numberstyle=\tiny, numbersep=5pt %, stepnumber=2, 
%         % literate={->}{{$\rightarrow$}}1 {<->}{{$\leftarrow$}}1
% }

% This Deep Tex Voodoo is from
%   <http://www.latex-community.org/forum/viewtopic.php?f=5&t=2072>
% It's purpose is to make \lstinline normal size, without affecting
% \lstinputlisting.  It seems to work but I have no idea how or why,
% and I rather hope never to learn.
\makeatletter
\lst@AddToHook{TextStyle}{\let\lst@basicstyle\ttfamily\normalsize}
\makeatother

\newcommand{\Figref}[1]{Figure \ref{#1}}
\newcommand{\Secref}[1]{Section \ref{#1}}
\newcommand{\x}[1]{\lstinline{#1}}

\input{formaldefs.tex}
\begin{document}

\conferenceinfo{SIGBOVIK '13}{Pittsburgh, PA, USA}
\copyrightyear{2013}
\copyrightdata{[to be supplied]}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{
  The Randomly-Scoped Lambda Calculus
}
% \subtitle{Subtitle Text, if any}

\authorinfo{Ben Blum}{}{bblum@cs.cmu.edu}

%\authorinfo{
%  Nicholas D. Matsakis
%  \and Brian Anderson
%  \and Ben Blum \\
%  \and Tim Chevalier 
%  \and Graydon Hoare
%  \and Patrick Walton
%  \and David Herman}
%           {Mozilla Research}
%           {\{nmatsakis, banderson\}@mozilla.com,
%             bblum@cs.cmu.edu, \\
%             \{tchevalier, graydon, pwalton, dherman\}@mozilla.com}

\maketitle

\begin{abstract}

We are inspired by the following travesty, courtesy of Python:
\begin{verbatim}
>>> x
NameError: name 'x' is not defined
>>> [x for x in 1,2,3]
[1, 2, 3]
>>> x
3
>>>
\end{verbatim}

The question of {\em scope} pertains to the set of rules governing which values a program's variables refer to during evaluation.
Prior work has proposed two main approaches: {\em static} (or {\em lexical}) {\em scope}, in which variable bindings are resolved through source code analysis, and {\em dynamic scope}, in which bindings are resolved at run-time using a stack of activation records.

In this work, we present {\em random scope}, a new technique for scoping, as an alternative to the above two.
Random scope affords programmers the flexibility to refer to multiple different binding sites simultaneously; for example, applying the term $(\lambda x. (\lambda x. x))$ to two arguments could evaluate to either the first or the second one.
We ask what kind of evaluation semantics it would take to make this work, and then pretty much stumble onwards from there trying to make sense of the whole damn thing.

We define the Randomly-Scoped Lambda Calculus, and show how it can evaluate random natural numbers as well as more popular deterministic functionalities such as factorial and fibonacci.

\end{abstract}

% \category{D.3.3}{Language Constructs and Features}{Data types and structures}

\keywords
static scope, dynamic scope, lower is better

\section{Introduction}

\bibliographystyle{alpha}
\bibliography{paper}

\end{document}