### 2013-03-21 tracks/tracks.tex
Update tracks to 2013

\documentclass[letter]{article}
\special{papersize=8.5in,11in}
\usepackage{fullpage}

\usepackage[lf]{electrum}
\usepackage[T1]{fontenc}

\newcommand\track[2]{\begin{center} {\Large \bf \em \sc Track #1} \vspace{0.2in}

{\huge \bf #2} \vspace{0.15in}
\end{center}
\thispagestyle{empty}}

\newcommand\asdf[2]{\vspace{0.15in} \item {\Large \bf \sf #1}

{\large \em #2}}
\newcommand\keywords[1]{

{\small {\bf Keywords:} #1}}

\newcommand\intermission[2]{}
%{\vspace{4em}\begin{tabular}{rl}
%	\Large \sc Intermission: & \Large \bf \sf #1 \\
%				 & \keywords{#2}
%\end{tabular}}
\begin{document}

\track{1}{Contemporary Issues in Politics and Law}

\begin{enumerate}
\asdf{SIGBOVIK License Agreement (excerpt)}
{Ix Tchow, Ph.D.}
\asdf{Redistributive Version Control Systems}
{Karlo Angiuli and Frederick Engels}
\keywords{redistribution, communism, version control}
\asdf{Psychology in Response to Presidential Poles}
{Timothy Broman}
\keywords{statistics, politics, election, cognitive dissonance, confirmation bias}
\end{enumerate}
\intermission{P-Rotip}{dreck, doggerel, puns}

\newpage

\track{2}{Harry Bovik's Research Calculus}

\begin{enumerate}
\asdf{The Randomly-Scoped Lambda Calculus}
{Ben Blum}
\keywords{static scope, dynamic scope, lower is better}
\asdf{Recovered Mathematical Journal}
{James McCann}
\asdf{MacLeod Computing: A New Paradigm for Immortal Distribution}
{Taus Brock-Nannestad, Gian Perrone, and Dr. Tom Murphy VII Ph.D.}
\keywords{cloude, compute, scottt}
\asdf{The ($\infty$,1)-Accidentopos Model of Unintentional Type Theory \\(extended abstract)}
{Carlo Angiuli}
\keywords{mistaken identity, accidentopos, object misclassifiers}
\end{enumerate}
\intermission{Pr-Otip}{art, artifice, artificial}

\newpage

\track{3}{Artificial Stupidity}

\begin{enumerate}
\asdf{You Only Learn Once: A Stochastically weighted AGGRegation Approach to Online Regret Minimization}
{danny ``vato loco'' m4turana and d4vid ``the H is silent'' f0uhey}
\keywords{yolo, swag, swagger, dat, bound, machine learning, ICML, brogrammerproblems, noregrets, belieber}
\asdf{I Lost The Game, and So Will You: Implications of Mindvirus Circulation in a Post-Singularity World}
{Shomir Wilson}
\keywords{singularity, game, experiment}
\asdf{Fandomized Algorithms and Fandom Number Generation}
{Lindsey Kuper and Alex Rudnick}
\keywords{fandomized algorithms, fandom numbers, fentropy}
\end{enumerate}
\intermission{pro-tip}{pro, tip, cycle}

\newpage

\track{4}{Computer Vision(aries)}

\begin{enumerate}
\asdf{Optimal Coupling and Gaybies}
{Nicolas Feltman}
\keywords{coupling, gaybies, neil patrick harris}
\asdf{Cat Basis Purrsuit}
{Daniel Caturana and David Furry}
\keywords{Meow, mew, miao, compressed sensing}
\asdf{A Spectral Approach to Ghost Detection}
{Daniel Maturana and David Fouhey}
\keywords{spectral and astral methods, trans-dimensional group lasso, ghosts, occult, hauntology, crystal energy, supernatural, paranormal}
\end{enumerate}
\intermission{prot-ip}{bad, twitter, joke}

\newpage

\track{5}{Productivity and Meta-Productivity \\ \large (or ``Synergistic Hyperparadigmatism'', as they say in the literature)}
\begin{enumerate}
\asdf{Really Amazing New Idea}
{Basquet Kase, Nought Job, and Ayn San Ety}
\keywords{ideas, doggerel, dreck}
\asdf{METHOD AND APPARATUS FOR PRESSING SPACEBAR}
{Joshua A. Wise and Jacob D. Potter}
\keywords{patent, spacebar, microcontrollers}
\asdf{METHOD AND APPARATUS FOR PUSHING SPACEBAR}
{Joshua A. Wise and Jacob D. Potter}
\keywords{patent, lego, spacebar}
\asdf{Find a Separating Hyperplane With This One Weird Kernel Trick \\(sponsored contribution)}
{Daniel Maturana and David F. Fouhey}
\keywords{kernel methods, fast money, overfitting, bayesian monopoly}
\asdf{On $n$-Dimensional Polytope Schemes}
{David F. Fouhey and Daniel Maturana}
\keywords{not pyramid scheme, polytope schemes, get rich fast, algebraic geometry, elimination ideals}
\asdf{DUI: A Fast Probabilistic Paper Evaluation Tool}
{Ivan Ruchkin and Ashwini Rao}
\asdf{Paper and Pencil: a Lightweight WYSIWYG Typesetting System}
{Paul Stansifer}
\end{enumerate}
\intermission{proti-p}{joke, got, old}

\newpage

\track{6}{Time Travel, Space Travel, and Other Fun Games for Children}
\begin{enumerate}
\asdf{A Proposal for Overhead-Free Dependency Management with Temporally Distributed Virtualization}
{Peter chapman, Deby Katz, and Stefan Muller}
\keywords{virtual time machine, time machine, virtualization}
\asdf{The $n$-People $k$-Bikes Problem}
{Lancer \r{A}ngstr\"{o}m}
\keywords{bikes, bixe, biologically-impelled konveyance, motion planning, contiguous allocation algorithms}
\asdf{The Problem of Heads of a Fighting Force from Long Ago}
{Leslie Lamport, Robert Shostak, and Marshall Pease}
\keywords{groups of brains of computers, talking in groups, accepting faults, friends agreeing with themselves, people who go to space, agreeing-group for computer making, ten hundred most used words, sorry to the people who wrote this paper the first time}
\asdf{duoludo: a game whose purpose is games}
{David Renshaw}
\keywords{game, games, gameses}
\asdf{The First Level of Super Mario Bros. is Easy with Lexicographic Orderings and Time Travel {\normalsize\dots after that it gets a little tricky.}}
{Dr. Tom Murphy VII Ph.D.*}
\keywords{computational super mario brothers, memory inspection, lexicographic induction, networked entertainment systems, pit-jumping}
\end{enumerate}

\end{document}