### 2014-03-19 proceedings/titlepage.tex

% NB. xelatex this file.

\documentclass{article}
\usepackage{fontspec}

\begin{document}

\setmainfont{Rumsey}
\setsansfont{Birbaslo}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

\begin{titlepage}
\begin{center}

% Upper part of the page. The '~' is needed because \\
% only works if a paragraph has started.
\textsc{\LARGE The Association for Computational Heresy}\\[0.5cm]

\textsc{\large presents}\\[1cm]

\textsc{\Large A Record of the Proceedings of}\\[0.5cm]
% Title
\HRule \\[0.4cm]
{ \Huge \sf SIGBOVIK 2014 \\[0.4cm] }

\HRule \\[2cm]

% Author and supervisor

The eight annual intercalary robot dance party in celebration

of workshop on symposium about Harry Q. Bovik's $2^6$th birthday

\vfill

% Bottom of the page
{\Large Carnegie Mellon University\\[0.25cm]
Pittsburgh, PA\\[0.3cm]
April 1, 2014}

\end{center}
\end{titlepage}

\end{document}