### 2014-03-19 proceedings/tracks.tex

\documentclass[letter]{article}
\special{papersize=8.5in,11in}
\usepackage{fullpage}

\newcommand\track[2]{\center {\Large \bf \em \sc Track #1} \vspace{0.2in}

{\huge \bf #2} \vspace{0.15in}
\thispagestyle{empty}}

\newcommand\asdf[2]{\vspace{0.15in} \item {\Large \bf \sf #1}

{\large \em #2}}
\newcommand\keywords[1]{

{\small {\bf Keywords:} #1}}

\begin{document}

\track{1}{Official 2012 Apocalypse Track}

\begin{enumerate}
\asdf{An Eschatological Approach: The Four Horsemen of Computer Science}
{The Reverend Nicolas A. Feltman and Joan E. Chao, Mother Supremum}
\keywords{Eschatology, Horsies, Indulgences}
\asdf{World Domination Through the Use of a Graphical Reduction of the Six Degrees of Separation Concept with Potential Robot Design for Mode of Implementation}
{L. S. Berg and P. J. Barton}
\keywords{Six Degrees of Separation, Kevin Bacon, Turtle Power, World Domination, Bad Cartoon Jokes, Topology, Reduction, Depression, Graphical Model Abuse}
\end{enumerate}

\newpage

\track{2}{Comestibility theory}

\begin{enumerate}
\asdf{Food for Thought: Dining Philosophers}{}
\keywords{hunger-driven devices, morsels, monads}
\asdf{Higher-Order Generalized Algebraic Pizzas}
{Brandon Bohrer and Samir Jindel}
\keywords{Curry-Pepper Isomorphism, Grease Monad, Monoids in the Category of Delicious, $A^\flat$ Mixolydian}
\vspace{-0.1in}
\asdf{Lollipops and Lemma Drops: the sweet, sweet logic of candy}
{William Lovas}
\asdf{Algorithms for k/n Power-Hours}
{Ben Blum, Dr. William Lovas, Ph.D., Chris Martens, and Dr. Tom Murphy VII, Ph.D.}
\keywords{alcohol in computer science, algorithms for the real world, chemically-assisted reasoning, drinking game theory, hyper-driven devices}
\end{enumerate}

\newpage

\track{3}{Brought to you by the letter\dots}

\begin{enumerate}
\setcounter{enumi}{-1}
\asdf{TBD}
{Taus Brock-Nannestad and Gian Perrone}
\keywords{all is well here, send money, love to you and yours}
\asdf{The Letter}
{Frederick J. Mesp\"{u}tchy}
\asdf{Proof of P = NP}
{Samir Jindel and Brandon Bohrer}
\asdf{An Epistolary Reconstruction of the Curry-Howard Correspondence}
{Ben Blum and Michael Sullivan}
\keywords{Corre-Howard-Spondence, Simply Typed Lambada Calculus, Supernatural Deduction}
\asdf{The Kardashian Kernel}
{David F. Fouhey and Daniel Maturana}
\keywords{Kardashian, Kim, Kourtney, Khloe, Kernel, Kuadratic, Konvex, Koncave, Krylov, Kronecker, Kolmogorov, Karush-Kuhn-Tucker, K-Means, K-Armed-Bandit, Kohonen, Karhunen-Loeve, Kriging, Kalman, Kinect, Kovariance, Kurse-of-dimensionality, Kurvature, K-Nearest-Neighbor}
\end{enumerate}

\newpage

\track{4}{Did you bring enough to reshare \\ with the class?}

\begin{enumerate}
\setcounter{enumi}{-1}
\asdf{Implications of Constrained Thought Expression Achieved via a One Hundred-forty Character Message Limitation Applied to Complex Social Netwo}
{Nathan Brooks and Tommy Liu}
\keywords{\#socialmedia, \#sm, \#140}
\asdf{The Spineless Tagless Tweet Machine: Distributed Cloud-Based Social Crowdsourced Lazy Graph Reduction on the Web 2.0}
{Michael Sullivan}
\asdf{SIGBOVIK 2012 Take-Home Midterm Examination}
{James McCann}
\asdf{The National Month Of Pushing Spacebar}
{Tom Murphy VII}
\keywords{C++harles D:{\textbackslash}ickens, fanfic, repetitive stress injury, 2012: A Space Odyssey}
\asdf{What Most Medical Students Know About Computer Science}
{Brian R. Hirshman}
\keywords{computer science, electronic medical records, all-knowing doctors, important stuff, take two kilobytes and see me in the morning, dude we're too busy learning other things}
\end{enumerate}

\newpage

\track{5}{Programming languages research and \\ other games for children}
\begin{enumerate}
\asdf{Modeling Perceived Cuteness}
{Nathan Brooks and Evelyn Yarzebinski}
\keywords{cute, puppy, kitten, turtle, spider, snake, dog, cat, gecko, hamster}
\asdf{i-PHONE app stor : where is my pants}
{Dr. Tom Murphy VII, Ph.D. and Dr. Abigale G. Lade, Ph.D.}
\keywords{pants, where, is my, juvenilia}
\asdf{An Extensible Platform for Upwards and Sidewards Mobility}
{David Renshaw}
\asdf{A modern optimizing compiler for a dynamically typed programming \\ language: Standard ML of New Jersey (preliminary report)}
{Ian Zerny}
\asdf{Programming Language Checklist}
{Colin McMillen, Jason Reed, and Elly Jones}
\end{enumerate}

\newpage

\track{6}{Protect yo' shit}
\begin{enumerate}
\asdf{Cryptographically-Strong Hashing with Physics}
{James McCann and Fake Otherauthor}
\asdf{The Physics-Based Hashes of Twelve Really Good Ideas}
{Patent Troll}
\asdf{A modest proposal for the purity of programming}
{Robert J. Simmons and Tom Murphy VII}
\keywords{the next 700 terrible programming languages, hyper-driven devices, types}
\asdf{Preparation-Hom as the ideal completion of a Hemorrhoidal category}
{Sitsin C{\o}mfort and Minr{\o}v G{\o}rondt}
\keywords{hemorrhoids, fistulads, co-fistulads, push outs}
\asdf{A Randomized Commit Protocol for Adversarial - Nay, Supervillain - \\ Workloads}
{Ben Blum}
\keywords{two-phase, two-face, randomized algorithms, acid}
\asdf{Address space content randomization: exploit mitigation through data \\ randomization}
{Carlo Angiuli}
\keywords{security, optimization, nondeterministic memory, lower is better}
\end{enumerate}

\end{document}