### 2014-03-20 review-2014/review-5.tex
finish mostly tracks and toc

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{2}{*}{\includegraphics[width=1in]{ach.png}}\\
& \textbf{\Huge{SIGBOVIK 2014 Paper Review}} \\ &\\
& \LARGE{Paper 5: Yet Another Application of Our Pet Technique} \\
&\\
\hline
\end{tabular}}
\vspace{2em}

{\large\bf
\begin{tabular}{l}
Robert Marsh, King Under the Mountain \\
Rating: 1 (strong reject) \\
Confidence: absolute/4 \\
\end{tabular}}
\vspace{1em}

There are several problems with this paper.
The modeling of vehicles is rather poor - at the very least their acceleration in the road's reference frame is limited by relativistic effects.
Furthermore, modern vehicles are curvy, not dumb and boxy (except the Nissan Cube, which isdumb, boxy, and curvy, truly a marvel of modern engineering) and thus would be best modeled as spheres.
In addition, the alternatives they test against don't include the current industry standard dime-an-hour hick, suggesting that they were cherrypicked to appear more favorable.
And besides, they didn't cite \emph{anything} I did, so I don't see why I should let them get published.

\end{document}