### 2014-03-20 proceedings/tracks.tex

\documentclass[letter]{article}
\special{papersize=8.5in,11in}
\usepackage{fullpage}
\usepackage{fontspec}
\usepackage{amsfonts}

\setmainfont{Rumsey}
\setsansfont{TragicSans}

\newcommand\track[2]{\center {\Large \bf \em \sc Track #1} \vspace{0.2in}

{\huge \bf #2} \vspace{0.15in}
\thispagestyle{empty}}

\newcommand\paper[2]{\vspace{0.15in} \item {\Large \bf \sf #1}

{\large \em \sf #2}}
\newcommand\keywords[1]{

{\small {\bf Keywords:} #1}}

\begin{document}

%\track{8}{Take The Quiz: Which Pointless Time-Wasting Amusement are You?}
\track{8}{How to Keep Adults Busy and Other Fun Games for Children}

\begin{enumerate}
\paper{How to Keep a Graduate Student Busy}
{Paul Stansifer}
\keywords{all, work, and, no, play, makes, paul, a, dull, boy}
\paper{New Results in $k/n$ Power-Hours}
{Dr. Tom Murphy VII Ph.D.}
\keywords{generalized binge drinking, maths, finite-state automata, abstract interpretation}
\paper{Linear Logic Free}
{Carlo Angiuli}
\keywords{linear logic, free to play, app store}
\paper{It Still Seems That Black Has Hope in These Extremely Unfair Variants of Chess}
{Dr. Tom Murphy VII Ph.D., Ben Blum, and Jim McCann, in italics ...... oh! .... nervous laughter .... no, that's not part of the name still. I stopped saying the name. Why are you still typing? This isn't the name any more. End quote.}
\keywords{chess, stratego, guess who, clue, candy crush saga, game-tree search, combinations}
\end{enumerate}

\newpage

%\track{88}{Epic Local News You'll Be Sad You Missed}
\track{88}{Concerning Local Affairs}

\begin{enumerate}
\paper{PARTIAL TRANSCRIPT OF DEAN'S SPEECH TO THE FACULTY OF THE MELLON COLLEGE OF SCIENCES}
{Clint Falamenos}
\keywords{news, lane center, college of sciences}
\paper{Analysis of the Effects of Substantial Lane Closure During Afternoon Peak Along a Heavily-Traveled Urban Arterial Choke-Point}
{Nicholas Fudala}
\keywords{knowledge, pagers, RATPAG}
\paper{Yet Another Application of Our Pet Technique}
{Chester Francis and Nicholas Fudala}
\keywords{complexity, multiple patents pending, nobel peace prize}
\paper{Please Don't Let Open House Destroy the Universe}
{Jenn Landefeld}
\keywords{open house, destruction of universe, herding cats}
\end{enumerate}

\newpage

%\track{888}{Four Unbelievable New Theoretical Perspectives}
\track{888}{New Theoretical Perspectives}

\begin{enumerate}
%\setcounter{enumi}{-1}
\paper{Heterotopy Type Theory: \\ A Defense of the Traditional Foundations of Mathematics}
{Thomas Cramer and Sam Yeager}
\keywords{heterotopy type theory, identity type, freedom of expression axiom}
\paper{The Dumping Lemma - Assessing Regularity}
{Naomi Saphra}
\keywords{computational scatology, formal languages, poop jokes}
\paper{Statistical Watch Optimization}
{Bryce Summers}
\keywords{Watch, clock, time, probability, statistics}
\paper{A Simple Category-Theoretic Understanding of Category-Theoretic Diagrams}
{Stefan Muller}
\keywords{category, diagram, functor, diagram diagram}
\end{enumerate}

\newpage

%\track{8,888}{Three Important Ways to Stay Safe This Research Season}
\track{8,888}{Stay Safe Out There}

\begin{enumerate}
%\setcounter{enumi}{-1}
\paper{Cryptographically-Sound Jokes}
{Jim McCann}
\keywords{jokes, sha, md5}
\paper{DollarCoin: A Cryptocurrency with Proof-of-Dollar}
{Patrick J. Xia}
\keywords{cryptocurrency,
bitcoin,
dollars,
bills,
gamechanger,
disrupt,
awesome,
supercool,
thebest,
hashtag}
\paper{Towards a Completely Autonomous Vehicle}
{Nicholas Fudala and Chester Francis}
\keywords{pain, suffering, frustration}
\end{enumerate}

\newpage

%\track{88,888}{Improve Your Own Research With These Five Weird Tricks}
\track{88,888}{Easier, Better, Faster, Stronger Research Methodologies}
\begin{enumerate}
\paper{Belief-Sustaining Inference}
{Alex Reinhart and Jerzy Wieczorek}
\keywords{uninformative likelihood,
statistics,
embarrassment,
cognitive dissonance,
inference}
\paper{Solutions to Ley Line Access in Occult Computing}
{Maija Mednieks}
\keywords{ley lines,
occult computing,
miskatonic university,
installing linux on a dead badger,
playing god}
\paper{Measuring Your $\mathbb{P}$-ness: Determining How Like a Probability Distribution a Given Mathematical Object Is}
{John T. Longwood}
\keywords{lowbrow humour,
probability,
statistics}
\paper{a clarification of terminology}
{David Renshaw}
\keywords{did you mean,
search instead for,
not to be confused with}
\paper{What, If Anything, is Epsilon?}
{Dr. Tom Murphy VII Ph.D.}
\keywords{computational archaeology,
epsilon,
very-small and medium-small numbers}
\end{enumerate}

\newpage

%\track{888,888}{These New Ways to Program Will Blow Your Mind}
\track{888,888}{Innovations in Natural and Unnatural Languages}
\begin{enumerate}
\paper{IRS-1040-Based Computing: The Untapped Computational Power of Bureaucracy}
{William Gunther and Brian Kell}
\keywords{unconventional computation,
taxes,
compilers,
bureaucracy}
\paper{Towards Fooling and Foiling the NSA: \\
Surveying the State of Steganographic Programming Languages}
{Wolfgang Richter and Debjani Biswas}
\keywords{steganography,
esoteric programming language,
esoteric,
programming languages,
obfuscation,
cryptography,
shit}
\paper{Pikachu, Domosaur, and other Monolexical Languages}
{Sarah Allen, Jesse Dodge, and Domosaur Domosaur}
\keywords{natural language processing,
regular languages,
translation}
\paper{Unit-Test-Based Programming}
{Miguel \'{A}. Lech\'{o}n}
\keywords{unit test,
automatic stubbornness,
holy grail}
\end{enumerate}

\end{document}