### 2014-03-20 proceedings/table-of-contents.tex
finish proceedings for this year & write knowledge in readme

\documentclass[letter]{article}
\special{papersize=8.5in,11in}
\usepackage{fullpage}
\usepackage{fontspec}
\usepackage{amsfonts}

\setmainfont{TragicSans}

\newcommand\track[2]{{\Large Track #1: #2}}

\newcommand\paper[2]{\item \sf #1 \dotfill #2}

\newcommand\keywords[1]{}

\begin{document}

\begin{center}
{\Huge You Won't Believe This Table of Contents}
\end{center}
\thispagestyle{empty}

\vspace{6em}

\track{8}{Take The Quiz: Which Pointless Time-Wasting Amusement are You?}
%\track{8}{How to Keep Adults Busy and Other Fun Games for Children}

\begin{enumerate}
\paper{How to Keep a Graduate Student Busy}
{3}
\keywords{all, work, and, no, play, makes, paul, a, dull, boy}
\paper{New Results in $k/n$ Power-Hours}
{5}
\keywords{generalized binge drinking, maths, finite-state automata, abstract interpretation}
\paper{Linear Logic Free}
{17}
\keywords{linear logic, free to play, app store}
\paper{It Still Seems That Black Has Hope in These Extremely Unfair Variants of Chess}
{21}
\keywords{chess, stratego, guess who, clue, candy crush saga, game-tree search, combinations}
\end{enumerate}

\vspace{4em}

\track{88}{Epic Local News Stories You'll Be Sad You Missed}
%\track{88}{Concerning Local Affairs}

\begin{enumerate}
\paper{PARTIAL TRANSCRIPT OF DEAN'S SPEECH TO THE FACULTY OF THE MELLON COLLEGE OF SCIENCES}
{29}
\keywords{news, lane center, college of sciences}
\paper{Analysis of the Effects of Substantial Lane Closure During Afternoon Peak Along a Heavily-Traveled Urban Arterial Choke-Point}
{33}
\keywords{knowledge, pagers, RATPAG}
\paper{Yet Another Application of Our Pet Technique}
{35}
\keywords{complexity, multiple patents pending, nobel peace prize}
\paper{Please Don't Let Open House Destroy the Universe}
{41}
\keywords{open house, destruction of universe, herding cats}
\end{enumerate}

\vspace{4em}

\track{888}{Four Unbelievable New Theoretical Perspectives}
%\track{888}{New Theoretical Perspectives}

\begin{enumerate}
%\setcounter{enumi}{-1}
\paper{Heterotopy Type Theory: A Defense of the Traditional Foundations of Mathematics}
{47}
\keywords{heterotopy type theory, identity type, freedom of expression axiom}
\paper{The Dumping Lemma - Assessing Regularity}
{51}
\keywords{computational scatology, formal languages, poop jokes}
\paper{Statistical Watch Optimization}
{53}
\keywords{Watch, clock, time, probability, statistics}
\paper{A Simple Category-Theoretic Understanding of Category-Theoretic Diagrams}
{57}
\keywords{category, diagram, functor, diagram diagram}
\end{enumerate}

\vspace{4em}

\track{8,888}{The Three Most Important Ways to Stay Safe This Research Season}
%\track{8,888}{Stay Safe Out There}

\begin{enumerate}
%\setcounter{enumi}{-1}
\paper{Cryptographically-Sound Jokes}
{65}
\keywords{jokes, sha, md5}
\paper{DollarCoin: A Cryptocurrency with Proof-of-Dollar}
{67}
\keywords{cryptocurrency,
bitcoin,
dollars,
bills,
gamechanger,
disrupt,
awesome,
supercool,
thebest,
hashtag}
\paper{Towards a Completely Autonomous Vehicle}
{71}
\keywords{pain, suffering, frustration}
\end{enumerate}

\vspace{4em}

\track{88,888}{Improve Your Own Research With These Five Weird Tricks}
%\track{88,888}{Easier, Better, Faster, Stronger Research Methodologies}
\begin{enumerate}
\paper{Belief-Sustaining Inference}
{77}
\keywords{uninformative likelihood,
statistics,
embarrassment,
cognitive dissonance,
inference}
\paper{Solutions to Ley Line Access in Occult Computing}
{83}
\keywords{ley lines,
occult computing,
miskatonic university,
installing linux on a dead badger,
playing god}
\paper{Measuring Your $\mathbb{P}$-ness: \\ Determining How Like a Probability Distribution a Given Mathematical Object Is}
{87}
\keywords{lowbrow humour,
probability,
statistics}
\paper{a clarification of terminology}
{91}
\keywords{did you mean,
search instead for,
not to be confused with}
\paper{What, If Anything, is Epsilon?}
{93}
\keywords{computational archaeology,
epsilon,
very-small and medium-small numbers}
\end{enumerate}

\vspace{4em}

\track{888,888}{These New Ways to Program Will Blow Your Mind}
%\track{888,888}{Novel Programming Methodologies}
\begin{enumerate}
\paper{IRS-1040-Based Computing: The Untapped Computational Power of Bureaucracy}
{101}
\keywords{unconventional computation,
taxes,
compilers,
bureaucracy}
\paper{Towards Fooling and Foiling the NSA: \\
Surveying the State of Steganographic Programming Languages}
{103}
\keywords{steganography,
esoteric programming language,
esoteric,
programming languages,
obfuscation,
cryptography,
shit}
\paper{Pikachu, Domosaur, and other Monolexical Languages}
{109}
\keywords{natural language processing,
regular languages,
translation}
\paper{Unit-Test-Based Programming}
{115}
\keywords{unit test,
automatic stubbornness,
holy grail}
\end{enumerate}
\thispagestyle{empty}

\end{document}