### 2017-03-14 itg/2/paper.tex
itg paper add a bunch of tables

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

%\documentclass[preprint]{sigplanconf}
\documentclass[10pt]{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{yfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
%\usepackage{mathpartir}
\usepackage{hyperref}
\usepackage{url}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{harmony}
\usepackage{marvosym}
\usepackage{multirow}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utopia]{mathdesign}
\usepackage{natbib}

% ____________________________________________________________
% Listings Package Configuration
% \usepackage[scaled]{beramono}

%\renewcommand*\ttdefault{txtt}
\usepackage[T1]{fontenc}

% This Deep Tex Voodoo is from
%   <http://www.latex-community.org/forum/viewtopic.php?f=5&t=2072>
% It's purpose is to make \lstinline normal size, without affecting
% \lstinputlisting.  It seems to work but I have no idea how or why,
% and I rather hope never to learn.
%\makeatletter
%\lst@AddToHook{TextStyle}{\let\lst@basicstyle\ttfamily\normalsize}
%\makeatother

\begin{document}

\conferenceinfo{SIGBOVIK '17}{Pittsburgh, PA, USA}
\copyrightyear{2017}
\copyrightdata{}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{
A Boring Follow-Up Paper to \\
``Which ITG Stepcharts are Turniest?'' \\
Titled, ``Which ITG Stepcharts are Crossoveriest and/or Footswitchiest?'' \\
%That `Boring' Stuff Was Part of the Title, BTW. \\
%So was that. And that, and this too. \\
%You got it all, right? \\
%Or Just, ``More Boring Crap about ITG'', for Short. \\
%Oh, That Was Also Part of the Title.
}
% \subtitle{\em The Randomly-Scoped Lambda Calculus}
% \subtitle{Subtitle Text, if any}

\authorinfo{Ben Blum}{}{bblum@cs.cmu.edu}

\maketitle

\begin{abstract}
	%ITG is a popular dance game in which players step on arrows while listening to music. The arrow patterns, indicated by a {\em stepchart}, may range among any level of complexity and difficulty. Among the many factors contributing to a stepchart's difficulty is how much the player must turn from side to side.
	%Other more obvious factors, such as raw speed, have been well studied in prior work.
	%This paper presents an analytic study of this {\em turniness} factor.
	%We study the turniness of many existing stepcharts, and present a novel (but unsurprising) approach to automatically generating maximally (or minimally) turny charts.
	%Among real-world songs, we find stepcharts with overall turniness ranging from 0\% to 81.33\% of the theoretical maximum.

In which I deliver on last year's promise of future work.

\end{abstract}

\category{D.D.R.}{Exercise and Fitness}{Arcade Dance Games}

\keywords
in, the, groove, 2

\section{Introduction}

Right, let's resume right where I left off in my last paper \cite{turniness}, shown in Figure~\ref{fig:you-stutid-fuckass}.
Unlike mainstream conferences, SIGBOVIK doesn't make me waste space repeating all the background material,
and I can just say go read that paper first and get back to me.
It's probably a lot funnier than this one anyway, which is gonna be sort of dry,
and really of interest only to other ITG players who already know what's going on.

\begin{figure}[h]
	\hspace{-2em}\includegraphics[angle=90,width=0.54\textwidth]{../paper.pdf}
	\caption{(okay twist your head to read this)}
	\label{fig:you-stutid-fuckass}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Flashback Scene}

Recall Table~\ref{tab:facing} from the last paper, in which I left undefined the facings for $LL$, $DD$, $UU$, and $RR$, the four footswitch footings.

\begin{table}[h]
	\begin{center}
	\begin{tabular}{cc|cccc}
		& & \multicolumn{4}{c}{Right foot} \\
		& & $\leftarrow$ & $\downarrow$ & $\uparrow$ & $\rightarrow$ \\
		\hline
		\multirow{4}{*}{Left foot}
		%                 L   D   U   R
		& $\leftarrow$  & - & $UR$ & $UL$ & $U$ \\
		& $\downarrow$  & $DL$* & - & $L$ & $UL$ \\
		& $\uparrow$    & $DR$* & $R$ & - & $UR$ \\
		& $\rightarrow$ & $D\dagger$ & $DR$* & $DL$* & - \\

	\end{tabular}
	\end{center}
	\caption{Facing directions. ``Crossover'' facings are marked (*), and the ``lateral'' facing is marked ($\dagger$). Note the appealing diagonal symmetry.}
	\label{tab:facing}
\end{table}

% dont forget
% http://dancedancerevolutionddr.wikia.com/wiki/Afronova_walk

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{pink}{RGB}{128,0,192}
\definecolor{orange}{RGB}{192,96,0}
\definecolor{olivegreen}{RGB}{0,127,32}
\definecolor{brickred}{RGB}{192,0,0}
\definecolor{commentblue}{RGB}{0,128,192}

\begin{figure*}[t]
\begin{center}
\begin{tabular}{l}
	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Step}~= \hilight{brickred}{L}~| \hilight{brickred}{D}~| \hilight{brickred}{U}~| \hilight{brickred}{R}~| \hilight{brickred}{Jump}~\hilight{orange}{deriving}~\hilight{olivegreen}{Eq}} \\
\texttt{} \\
\texttt{\hilight{orange}{data}~\hilight{olivegreen}{AnalysisState}~= \hilight{brickred}{S}~\{ steps :: \hilight{olivegreen}{Int}, xovers :: \hilight{olivegreen}{Int}, switches :: \hilight{olivegreen}{Int}, jacks :: \hilight{olivegreen}{Int},} \\
\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastStep :: \hilight{olivegreen}{Maybe}~\hilight{olivegreen}{Step}, doubleStep :: \hilight{olivegreen}{Bool}, lastFlip :: \hilight{olivegreen}{Bool},} \\
\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastFoot :: \hilight{olivegreen}{Bool}, stepsLR :: [\hilight{olivegreen}{Bool}] \}} \\
\texttt{} \\
\texttt{\hilight{pink}{commitStream}~:: \hilight{olivegreen}{AnalysisState}~-> \hilight{olivegreen}{AnalysisState}} \\
\texttt{\hilight{pink}{commitStream}~s = s \{ xovers~~ = xovers~~ s + \hilight{orange}{if}~f \hilight{orange}{then}~ns - nx \hilight{orange}{else}~nx, } \\
\texttt{~~~~~~~~~~~~~~~~~~~~ switches = switches s + \hilight{orange}{fromEnum}~(f == lastFlip s \&\& doubleStep s), } \\
\texttt{~~~~~~~~~~~~~~~~~~~~ jacks~~~~= jacks~~~~s + \hilight{orange}{fromEnum}~(f /= lastFlip s \&\& doubleStep s), } \\
	\texttt{~~~~~~~~~~~~~~~~~~~~ lastFlip = f, stepsLR = \hilight{brickred}{[]}~\}} \\
\texttt{~~~~\hilight{orange}{where}~ns = \hilight{orange}{length}~\$~stepsLR s} \\
\texttt{~~~~~~~~~~nx = \hilight{orange}{length}~\$ \hilight{orange}{filter not}~\$ stepsLR s} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- if more than half the L/R steps in this stream were crossed over,}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- then we got the footing backwards and need to flip the stream.}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- as a tiebreaker, flip if the chart is already more jacky than}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- footswitchy, i.e., if past streams flipped more often than not.}} \\
\texttt{~~~~~~~~~~f = nx * 2 > ns || nx * 2 == ns \&\& ((switches s > jacks s) == lastFlip s)} \\
\texttt{} \\
\texttt{\hilight{pink}{analyzeStep}~:: \hilight{olivegreen}{AnalysisState}~-> \hilight{olivegreen}{Step}~-> \hilight{olivegreen}{AnalysisState}} \\
\texttt{\hilight{pink}{analyzeStep}~s step} \\
\texttt{~~~~\hilight{commentblue}{-{}- a jump resets the footing, so the next step can be stepped with either}} \\
\texttt{~~~~\hilight{commentblue}{-{}- foot. commit the stream so far to treat it separately from what follows.}} \\
\texttt{~~~~\hilight{commentblue}{-{}- bracket-jumps are, of course, future work.}} \\
\texttt{~~~~| step == \hilight{brickred}{Jump}~= (\hilight{pink}{commitStream}~s) \{ lastStep = \hilight{brickred}{Nothing}, doubleStep = \hilight{brickred}{False}~\}} \\
\texttt{~~~~\hilight{commentblue}{-{}- two steps on the same arrow might be a jack, or might be a footswitch.}} \\
\texttt{~~~~\hilight{commentblue}{-{}- to figure out which, commit the stream so far, and begin a new stream}} \\
\texttt{~~~~\hilight{commentblue}{-{}- whose footing will retroactively determine how to foot this step.}} \\
\texttt{~~~~\hilight{commentblue}{-{}- also, unlike jumps, this step gets counted as part of the next stream.}} \\
\texttt{~~~~| lastStep s == \hilight{brickred}{Just}~step = stream (\hilight{pink}{commitStream}~s) \{ doubleStep = \hilight{brickred}{True}~\}} \\
\texttt{~~~~\hilight{commentblue}{-{}- a normal streamy step.}} \\
	\texttt{~~~~| \hilight{orange}{otherwise}~= stream s} \\
\texttt{~~~~\hilight{orange}{where}~foot = \hilight{orange}{not}~\$ lastFoot s} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- record whether we stepped on a matching or crossed-over L/R arrow.}} \\
\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{L}~steps = steps ++ [ft]} \\
\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{R}~steps = steps ++ [\hilight{orange}{not}~ft]} \\
\texttt{~~~~~~~~~~addStep ft \_ steps = steps \hilight{commentblue}{-{}- U/D don't help to determine L/R footing.}} \\
\texttt{~~~~~~~~~~stream s = s \{ steps = steps s + \hilight{brickred}{1}, lastStep = \hilight{brickred}{Just}~step, lastFoot = foot,} \\
\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ stepsLR = addStep foot step \$ stepsLR s \} } \\
\texttt{} \\
\texttt{\hilight{pink}{analyze}~:: [\hilight{olivegreen}{Step}] -> \hilight{olivegreen}{AnalysisState}} \\
	\texttt{\hilight{pink}{analyze}~= \hilight{pink}{commitStream}~. \hilight{orange}{foldl}~\hilight{pink}{analyzeStep}~(\hilight{brickred}{S 0 0 0 0}~\hilight{brickred}{Nothing}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{[]}) } \\
\end{tabular}
\end{center}
\caption{it's pseudocode lol}
\label{fig:pseudocode-lol}
\end{figure*}

\section{Existing Chart Analysis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section{A Bit More on Turniness}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Experimental Results}

% TODO: Put a * on charts where you're going to put a pict of mario man
% TODO: Put a dagger on footswitch charts where the footing is a clever reinterpretation, rather than intended
% TODO: Make a special mention of sandstorm and put a pict of mario man

\begin{figure}[t]
	\begin{center}
		\small
	\begin{tabular}{r|l|l|c|c}
		\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
		\hline
		\multicolumn{5}{c}{\em no 8s- with $\ge$100 XOs} \\
		 9 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 107 & \em 8.7 \\
		10 & J-PARA SUPER MEGAMIX     & CuoReNeRo MeGaPacK    & \bf 141 & \em 3.9 \\
		11 & Somebody That I Used To Know & Best of Gazebo    & \bf 114 & \em 11.9 \\
		12 & Credens Justitiam        & Stuff B Likes         & \bf 136 & \em 12.4 \\
		13 & Banshee Strikes          & VocaJawnz             & \bf 152 & \em 10.2 \\
		14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
		15 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 101 & \em 4.2 \\
		\multicolumn{5}{c}{\em no 16s-20s with $\ge$100 XOs} \\
		21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
	\end{tabular}
	\end{center}
	\caption{Charts with the most total crossovers (XOs).}
\end{figure}
\begin{figure}[t]
	\begin{center}
		\small
	\begin{tabular}{r|l|l|c|c}
		\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
		\hline
		 4 & DAM DIRIRAM              & DDR 3rdMIX            & \em  30 & \bf 27.3 \\
		 5 & STRICTLY BUSINESS        & DDR 1st               & \em  33 & \bf 21.4 \\
		 6 & STRICTLY BUSINESS        & DDR 1st               & \em  38 & \bf 20.9 \\
		 7 & MOBO$\star$MOGA          & DDR EXTREME           & \em  37 & \bf 17.3 \\
		 8 & PARANOiA                 & DDR 1st               & \em  58 & \bf 19.6 \\
		 9 & Dazzlin Darlin           & r21twins              & \em  76 & \bf 22.0 \\
		10 & Enchanted Journey        & ITG Rebirth           & \em  78 & \bf 19.3 \\
		11 & Lune Noir                & r21freak Friendship   & \em  58 & \bf 13.4 \\
		12 & Winnipeg is Fucking Over & best of r21freak ii   & \em 122 & \bf 14.3 \\
		13 & The Sampling Paradise    & The Paradise Sampler  & \em 139 & \bf 15.6 \\
		14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
		\multicolumn{5}{c}{\em no 15s-20s with $\ge$10\% XO} \\
		21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
	\end{tabular}
	\end{center}
	\caption{Charts with the highest percentage of crossovers among total steps.}
\end{figure}
\begin{figure}[t]
	\begin{center}
		\small
	\begin{tabular}{r|l|l|c|c}
		\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
		\hline
		\multicolumn{5}{c}{\em no 5s- with $\ge$50 FSs} \\
		 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
		\multicolumn{5}{c}{\em no 7s-9s with $\ge$50 FSs} \\
		10 & Eat 'Em Up!      & Mute Sims 5           & \bf  54 & \em  9.0 \\
		11 & Nemeton          & The Legend of Zim 4   & \bf  97 & \em 10.3 \\
		12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
		13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
		14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
		15 & Danse Macabre    & Aoreo's Ariginals     & \bf 201 & \em 12.2 \\
		16 & Weird Science    & Stamina Showcase      & \bf  61 & \em  1.8 \\
		\multicolumn{5}{c}{\em no 17s+ with $\ge$50 FSs} \\
	\end{tabular}
	\end{center}
	\caption{Charts with the most total footswitches (FSs).}
\end{figure}
\begin{figure}[t]
	\begin{center}
		\small
	\begin{tabular}{r|l|l|c|c}
		\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
		\hline
		%\multicolumn{5}{c}{\em no <6s with >50 FSs} \\
		 3 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  26 & \bf 15.3 \\
		 4 & MAKE IT BETTER   & DDR 2ndMIX            & \em  26 & \bf 20.0 \\
		\multicolumn{5}{c}{\em no 5s with $\ge$10\% FS} \\
		 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
		 7 & 5.1.1            & DDR MAX               & \em  13 & \bf 12.9 \\
		\multicolumn{5}{c}{\em no 8s with $\ge$10\% FS} \\
		 9 & PARANOiA KCET    & DDR 2ndMIX            & \em  33 & \bf 10.2 \\
		10 & Delhi Ill        & Mute Sims 6           & \em  40 & \bf 11.6 \\
		11 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  75 & \bf 11.8 \\
		12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
		13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
		14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
		15 & Flames of the Sky & fort rapids vii      & \em 172 & \bf 16.5 \\
		\multicolumn{5}{c}{\em no 16s+ with $\ge$10\% FS} \\
	\end{tabular}
	\end{center}
	\caption{Charts with the highest percentage of footswitches among total steps.}
\end{figure}
\begin{figure}[t]
	\begin{center}
		\small
	\begin{tabular}{r|l|l|c|c}
		\bf Ft. & \bf Name & \bf Pack & \bf \#XF & \bf XF\% \\
		\hline
		\multicolumn{5}{c}{\em no 8s- with $\ge$12 XFs} \\
		 9 & Dr. Boom-Bombay   & fort rapids vii            & \bf 18 & 4.0 \\
		10 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.5 \\
		11 & Heart Shooter     & VocaJawnz                  & \bf 44 & 5.1 \\
		12 & Beautiful Synergy & Keyboard Collaboration III & \bf 14 & 3.2 \\
		13 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.2 \\
		14 & Fancy Footwork    & Cirque du Zeppelin         & \bf 40 & 4.2 \\
		\multicolumn{5}{c}{\em no 14s+ with $\ge$12 XFs} \\
	\end{tabular}
	\end{center}
	\caption{Charts with the most crossover footswitches (XFs). (Here I chose 12 as the cut-off because below that are a bunch of false-positives from DDR.)}
\end{figure}

\section{Conclusion}

Please accept our paper.
We worked hard on it.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrvnat}
\bibliography{citations}

\end{document}