### 2023-03-27 seitan/seitan.tex
more important changes

\documentclass[12pt]{article}

\usepackage[charter]{mathdesign}

\setlength{\parskip}{0.5em}

\newcommand\chapter[2]{{\thispagestyle{empty} \Large \sc #1. \quad #2

\vspace{1.5em}}}

\begin{document}

\chapter{1}{Homesick}

It's summer 2020, and you are the city of Seattle PNW USA.

Within you, in upper Queen Anne, lives this story's protagonist.
They recently moved into you, far too recently to have deep enough social roots for the epidemiological circumstances of the year in question.
Queen Anne is a beautiful neighborhood,
full of cherry blossom trees stunning in the spring,
radio towers that one can stand at the base of and crane their neck to see the top in a dizzying twist of perspective
(except during wildfire season, when, incidentally, one's newly-acquired KN95s find a secondary use),
and artfully manicured flower gardens afore multimillion-dollar homes,
which make our protagonist feel keenly alienated in their 800sqft apartment.
They cope with quarantine, in part, by attending the now-virtual ThursDz,
%, now occurring virtually,
and they begin to long for a particular food of home.
%
Said home, Pittsburgh Appalachia USA, is famous for many foods,
but they long in particular for a less famous one: the humble seitan wing.

You, city of Seattle, do not know what a seitan wing is.

Seitan wings is quite simple, in concept: Take chicken wings, ideally buffalo-sauced with a side of blue cheese dip, and replace the meat with seasoned dough of vital wheat gluten.
You are no stranger to seitan itself, of course, boasting a bustling vegan scene, including at least two Thai joints whose menus' every single meat entr\'ee is actually seitan.
But -- as our protagonist will only later theorize -- the preparation with buffalo sauce is an innovation unique to a city where the hipster foodie and blue-collar bar food scenes intersect.
Presently, said theory yet an inkling in their mind, they search countless delivery menus.
Their final attempt, upon recommendation of a CMU acquaintance they thought understood their struggle, a place from Capitol Hill;
whose seitan ``appetizer'' turns out to be a long, tough stick most akin to a Sl*m Jim, tossed in a thick coating of bread crumbs more voluminous than the inner stick jimself.

Resignedly, the protagonist asks Google Trends what, exactly, the fuck. Google-{\em sama} replies with a color-coded map of interest by US state: only Pennsylvania, New York, and Colorado are shaded at all.

You, city of Seattle, are invited to pause reading, issue the same Google Trends query, and reflect upon your sins.

\newpage \thispagestyle{empty}

\chapter{2}{Secrets}

It's fall 2020, and you are Google's search results for ``seitan wings recipe''.
You know what a seitan wing is.

You profess to contain one and a half million results.
The protagonist doesn't have quite that much time on their hands,
but they do have a lot, so they try (or at least read) everything on your first page.
You don't seem to agree with yourself very much, they notice:
you sometimes suggest breading the wing (which the protagonist knows to be wrong),
you sometimes suggest BBQ or garlic sauce instead of buffalo (which is noncanonical, but fine),
but most bafflingly, you suggest as many different ways of cooking the seitan as there are recipes.
Most often, you suggest to boil it,
which the protagonist gamely tries,
all the while remembering a story they heard long ago about how their mom's dad was so bad a cook she once caught him boiling a steak.
It sucks.
You say they should boil it then bake it;
you say they should boil it then saut\'e it;
you say they should saut\'e it then boil it;
you say they should fry it;
you say they should not fry it, because it will leave the inside raw and gummy (this, for once, is true),
you say they should bake it;
you say they should bake it in broth;
you say they should bake it dry tossed in oil.

Among these, the last idea at least produces something both edible and texturally close to the crispy exterior the protagonist remembers, but it is only ``close'' like standing on the next tallest mountain peak and looking across the way at the one you {\em meant} to climb is ``close''.
Everything you say is touched by the spectre of health food writing,
because your think your reader must be into \textasciitilde{}vegan alternatives\textasciitilde{}.
Drowning in your million useless faces the protagonist knows one truth deep in their heart: Seitan wings is not a health food.

Months pass, and the protagonist does not talk to you anymore.
They play Breath of the Wild in Japanese (which had long been a goal of theirs),
they play lots of gay VNs on itch dot io,
they practice Touhou music on the piano,
%they play Touhou,
and they play Netrunner worlds online over an enchanting coq au vin their elderly neighbor prepared for their nascent no-contact food relationship.

It's December 25, and thinking of you again, they pose your unanswered mystery to Tw*tter.
A Seattle friend replies, at long last, with the secret:

Steam, then fry.

\newpage \thispagestyle{empty}

\chapter{3}{Respite}

It's early June 2021, and you are M*d Mex Big Burrito Restaurant Group restaurant.
%
You serve the canonical implementation of seitan wings
in Shadyside Pittsburgh Appalachia USA.
If the NIST standards committee filed an entry for the food, they would surely enshrine one of yours.

The protagonist, vaccinated, visits your city. Their city.

You are one of the first stops they make
(friends in tow, of course),
a high honor in the book of a foodie ex-grad-student who knows the East End inside and out.
They have practiced Chapter 2's secret lost art, but something is still wrong,
and sitting at your table in the summer heat they realize what it is:
technique aside, even fully cooked, their dough yields a chewy, glutinous inside,
whereas your insides are perfectly soft;
one's teeth pierce the fried shell and sink effortlessly through to the other side
in a delightful textural contrast.
They do not know how to replicate it, and at this point are too invested in the narrative to just ask.

They stay for three weeks in a ThursDz friend's guest room, only a little jealous of his homeownership.
They tell him the preceding 2 chapters, and together they make what is known so far.
He likes it, but agrees it's too chewy.

They play board games; they attend art festivals; they share a ride up to the observatory on a borrowed bike nicer than their own.

They realize they need to move back.

\newpage \thispagestyle{empty}

\chapter{4}{Tenderness}

It's late June 2021, and you are the ``Seitan Reloaded'' recipe posted by author {\sf @simrob} on Blogging Website M*dium Dot Com.

With the last week of their month-long vax celebration trip,
the protagonist visits two other ThursDz friends in Raleigh.
They speedrun Celeste (a nervous cat who never shows affection to strangers, except nonbinary people).
Said friends are themselves a little like cats, visibly changed by their own quarantine experiences.
All three humans are eager to share whatever fruits said experiences have borne.
One of them is you.

You, Seitan Reloaded, are not seitan wings.
You are sausage, strongly spiced, in need of no sauce,
to be put on scrambled eggs or pizza or simply to be eaten unadorned and cold from the fridge.
Your author demonstrates you to the protagonist.
He puts half a cup of paprika into you.
He pours soy sauce into you, and keeps pouring -- the protagonist agog -- for impossibly long.
He kneads you,
and the protagonist remarks that they were afraid to knead their own because it already came out so tough,
and he remarks that the olive oil and the nooch inside you help disrupt gluten formation,
% so even with all the kneading it comes out pretty tender,
and they think ah, this might be the final piece of the puzzle.

June being Pride month,
the protagonist's employer attempts some trans-themed programming,
% (new from the generic corporate rainbows they usually do),
but they dramatically neglect the ``nothing about us without us'' test
and commit several terrible faux pas.
Trans employees complain in slack,
and the brass tone police them into silence.
One is nearly fired.
The protagonist, already having been silenced previously,
can only try to help behind the scenes,
and it's good they're traveling so they have someone to weakly say ``hold me'' to
while they work through their trauma response.
Like always in protests, the surface problems get fixed, and the rest are left to become scars in the sands of time -- almost as long as it takes to knead you.

With only a week in this visit, between that and all the video games the humans need to nerd over, there isn't time to synthesize your wisdom into the main quest.
Instead they fry up slices of you, toss you in buffalo sauce (as if you need it),
and produce, speaking extremely charitably, a Scientific Byproduct.

But cooking, baking, working matter with one's hands, is a wonderful grounding activity,
and your texture, your tenderness is perfect.

% Your texture is perfect.

\newpage \thispagestyle{empty}

\chapter{5}{Wings}

It's August 2021, and you are visiting your parents for the first time since quarantine in Santa Barbara Socal USA.

%It's the first time you've seen them since quarantine, and
You have a lot to catch up on.
Where there was border collie Louie, there is now border collie Junior.
You will never know if Louie could tell you smell different now.
You are teaching your mom to play Heaven's Vault,
your new favorite conlang-learning archaeology VN,
trying to tell her that the helplessness the game makes her feel is in fact diagetic,
and she (also a language nerd) is kinda actually getting it.
%
Your psychological association is still tenuous,
because your gender is new to them and your body is new to all three of you,
but your parents will get there later, beyond the scope of this work.

They take you on their traditional morning walk. % along the beach.
You always preferred the snow to the beach,
%but you come to understand why your mom does,
but you still relish the feeling of sand in your toes and saltwater lapping at your calves.
You had not felt the ocean on them since you made them yours.
You walk Junior and you get seafood together from a stall along the boardwalk
and suddenly the city of your birth feels like a bubble of immunity to that disaffected-millenial grief you carry everywhere else you go.

You tell them the preceding 4 chapters,
and they agree to share with you the honor of putting the pieces together,
and that is the story of how seitan wings were made correctly in Santa Barbara for the first time ever.

\vspace{15em}

\noindent You thank Ellen Scherb, Will Scott, William Lovas, Rob Simmons, Chris Martens, Eve Blum, David Blum, and Jenny Lin for their invaluable roles in the story.

\newpage \thispagestyle{empty}

\begin{center}
	% {\sc \Large ben blum's seitan wings}
	% {\sc \Large trans rights seitan wings}
	{\sc \large Quarantine seitan wings}
\end{center}

Based on Chapter 4's viewpoint character.
You will need a steamer and either an immersion blender or a lot of wrist stamina.
Makes 4-6 servings.

\vspace{1em}

\noindent \begin{tabular}{p{0.95\textwidth}}
	{\bf Ingredients (Wing; dry)} \\
	\hline
	{\em 14 oz} -- Vital wheat gluten \\
	{\em 1 oz} -- Nutritional yeast (or parmesan) \\
	{\em 3 Tbsp} -- Spices: rosemary, thyme, onion or garlic powder, and white pepper \\
	{\em 2 tsp} -- Salt \\
	\\
	{\bf Ingredients (Wing; wet)} \\
	\hline
	{\em 12 oz} -- Water (or chicken stock, for irony) \\
	{\em 6 oz} -- Tomato paste \\
	{\em 2 oz} -- Olive oil \\
	\\
	{\bf Ingredients (Buffalo)} \\
	\hline
	{\em 2 Tbsp} -- Butter (omit for vegans) \\
	{\em 2 Tbsp} -- Hot sauce (Frank's is canonical, but you prefer Cholula) \\
	\\
	{\bf Ingredients (Bluech)} \\
	\hline
	{\em 1/3 cup} -- Blue cheese (a nice, creamy, pungent one from Penn Mac) \\
	{\em 1/3 cup} -- Buttermilk (you can substitute but it really makes a difference) \\
	{\em Spoonful} -- Plain greek yogurt or sour cream, for consistency, optional \\
	\\
	{\bf Method} \\
	\hline
\setlength{\parskip}{0.5em}
	In two bowls, separately mix the dry and wet wingredients. Combine and knead for 5-10 minutes, until coherent but also tears easily. Add more liquid or gluten as necessary, a little at a time.

	Tear by hand (scraggly is good) into 1" chunks and steam for 30 minutes.

	Meanwhile, chongle the blue cheese dip: Reserve half the cheese, and immersion blend everything else. Add reserved cheese and mash together with a fork until pleasantly chunky.

	Melt butter and then mix in hot sauce. Do not microwave your hot sauce.

	Fry steamed wings in neutral oil at 350$^{\circ}$F for 3-5 minutes, until just crispy. Pat dry with a paper towel, toss in sauce, and serve over math puzzles.
\end{tabular}

\end{document}